Article Title

### 2013-12-02 plos.tex
updated the Simulation section

To illustrate key concepts, Figure \ref{fig:01} illustrates a regularization scheme based on \emph{covariance selection} \cite{Dempster:1972}. In covariance selection, the estimate is produced by fitting only an optimal subset of the coefficients of the \emph{precision matrix} (the inverse of the covariance matrix) while setting the remaining coefficients to zero. Scaling the precision matrix so that its diagonal is composed of all $-1$'s produces the matrix of partial correlations. Thus zeros in the precision matrix produce zero partial correlations between the corresponding pairs of neurons.  Panel \ref{fig:01}\,A depicts the sample correlation matrix of a period of somatic calcium signals of a group of 298 neurons in mouse visual cortex.  Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are much higher than typical firing rate correlations on shorter temporal scales. Panel \ref{fig:01}\,B magnifies the fragment of the sample correlation matrix outlined in \ref{fig:01}\,A whereas \ref{fig:01}\,C shows the corresponding fragment of the regularized estimate. The regularized estimate is produced from the same data by zeroing the optimal set of 31501 (71.2\%) of off-diagonal coefficients of the precision matrix and fitting the remaining coefficients. Panels \ref{fig:01}\,D and E show the same fragments of the respective estimates of the partial correlation matrix. The apparent similarity of the two estimates of the correlation matrix belies the utterly different partial correlation structure. The regularized estimate is less sensitive to sample noise in the data, is likely to be closer to truth in most cases than the unbiased estimate, which may be ascertained by cross-validation studies.

@@

For the purposes of this study, we adopted the \emph{negative normal log-likelihood loss}\footnote{
The loss function $\loss(C,\Sigma)$ is related to the multivariate log-likelihood function $L(\Sigma|C_0)$ as 
\begin{equation}
    L(\Sigma|C_0) = -\frac p 2 \ln(2\pi) -\frac p 2 \loss{\Sigma,C_0}
\end{equation}
}
defined as:
\begin{equation}\label{eq:loss}
    \loss{C,\Sigma} = \frac 1 p\left[\ln \det C + \Tr(C^{-1}\Sigma)\right]
\end{equation}

We normalized the loss function  by $\frac 1 p$ to make its value less sensitive population size (For independent neurons, the log likelihood scales linearly with $p$).

This choice is motivated by mathematical convenience. Other popular choices for the loss function are the Frobenius norm of the difference $C-\Sigma$ \cite{Ledoit:2004,Schafer:2005}, Stein's entropy loss, and quadratic loss \cite{James:1961,Fan:2008}.  We expect that the main conclusions of our study will not change qualitatively under other well behaved loss functions.

To make a loss function that assume zero at its minimum when $C=\Sigma$, we define \emph{excess loss} as
\begin{equation}\label{eq:excess-loss}
    \eloss{C,\Sigma} = \loss{C,\Sigma}-\loss{\Sigma,\Sigma}
\end{equation}

@@

As expected, estimators with matching low-dimensional structures typically outperformed the other estimators. There were two exceptions to this observation. For small sample sizes, before the data were sufficient to reveal the true correlation structure and to  allow the correct model to dominate, estimates with simpler targets often outperformed the more matching estimate.

With empirical data, the truth $\Sigma$ is not accessible and the relative values of loss must be estimated solely from the data. This may be accomplished through \emph{validation}.  In validation, the data are split into a training subset from which the estimate is obtained and a validation subset. We then compute the \emph{validation loss}

\begin{equation}\label{eq:validation-loss}
    \loss{C,C^\prime_0}
\end{equation}

$\loss{\cdot,\cdot}$ is additive in its second argument so that 
 \begin{equation}\label{eq:additivity}
 \loss{\hat\Sigma,X_1} + \loss{\hat\Sigma,X_2} \equiv \loss{\hat\Sigma,X_1+X_2}
 \end{equation}
Then \emph{validation loss}  
\begin{equation}\label{eq:validationLoss}
\hat \ell = \loss{\hat\Sigma,\hat\Sigma_0^\prime}
\end{equation}
is an unbiased estimate of the risk:
 \begin{equation}\label{eq:empiricalRisk}
\E[(\hat\Sigma, \hat\Sigma_0^\prime)] {\hat\ell} 
= \E[(\hat\Sigma, \hat\Sigma_0^\prime)]{\loss{\hat\Sigma,\hat\Sigma_0^\prime}}
= \E{\loss{\hat\Sigma,\E{\hat\Sigma_0^\prime}}}
= \E{\loss{\hat\Sigma,\Sigma}} = r
 \end{equation}
Therefore, estimators resulting in consistently lower validation loss can be inferred to produce estimates that are closer to truth than estimators with higher validation loss.

Other popular loss functions such as entropy loss do not comply with Eq.~\ref{eq:additivity} and their validation loss is not an unbiased estimate of risk.

>>>>

To illustrate key concepts, Figure \ref{fig:01} illustrates a regularization scheme based on \emph{covariance selection} \cite{Dempster:1972}. In covariance selection, the estimate is produced by fitting only an optimal subset of the coefficients of the \emph{precision matrix} (the inverse of the covariance matrix) while setting the remaining coefficients to zero. Scaling the precision matrix so that its diagonal is composed of all $-1$'s produces the matrix of partial correlations. Thus zeros in the precision matrix produce zero partial correlations between the corresponding pairs of neurons.  Panel \ref{fig:01}\,A depicts the sample correlation matrix of a period of somatic calcium signals of a group of 298 neurons in mouse visual cortex.  Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are much higher than typical firing rate correlations on shorter temporal scales. Panel \ref{fig:01}\,B magnifies the fragment of the sample correlation matrix outlined in \ref{fig:01}\,A whereas \ref{fig:01}\,C shows the corresponding fragment of the regularized estimate. The regularized estimate is produced from the same data by zeroing the optimal set of 31,501 (71.2\%) off-diagonal coefficients of the precision matrix and fitting the remaining coefficients. Panels \ref{fig:01}\,D and E show the same fragments of the respective estimates of the partial correlation matrix. The apparent similarity of the two estimates of the correlation matrix belies the utterly different partial correlation structure. The regularized estimate is less sensitive to sample noise in the data, is likely to be closer to truth in most cases than the unbiased estimate, which may be ascertained by cross-validation studies.

@@

In this study, we adopted \emph{negative normal log-likelihood loss}\footnote{
The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelihood function $L(\Sigma|C_0)$: 
\begin{equation*}
    L(\Sigma|C_0) = -\frac p 2 \ln(2\pi) -\frac p 2 \loss{\Sigma,C_0}
\end{equation*}
}
defined as
\begin{equation}\label{eq:loss}
    \loss{C,\Sigma} = \frac 1 p\left[\ln \det C + \Tr(C^{-1}\Sigma)\right]
\end{equation}

We normalized the loss function  by $\frac 1 p$ to make its value comparable across different population sizes.

This choice is motivated by mathematical convenience. Other popular choices for the loss function are the Frobenius norm of the difference $\|C-\Sigma\|_F$ \cite{Ledoit:2004,Schafer:2005}, Stein's entropy loss, and quadratic loss \cite{James:1961,Fan:2008}.  We expect that the main conclusions of our study will not change qualitatively under other well behaved loss functions.

To make the loss function to assume zero at its minimum when $C=\Sigma$, we define \emph{excess loss} as
\begin{equation}\label{eq:excess-loss}
    \eloss{C,\Sigma} = \loss{C,\Sigma}-\loss{\Sigma,\Sigma}
\end{equation}

@@

As expected, estimators with matching low-dimensional structures typically outperformed the other estimators. There were two exceptions to this observation. For small sample sizes, before the data were sufficient to reveal the true correlation structure and to  allow the correct model to dominate, estimates with simpler targets often outperformed the matching estimator.

When the ground truth $\Sigma$ is not accessible, loss can be estimated solely from the data through \emph{validation}.  In validation, an additional independent \emph{testing dataset} is used to compute an independent sample covariance estimate $C_0^\prime$.  Then \emph{validation loss} $\loss{C,C_0^\prime}$ can be used as an unbiased estimate\footnote
{
    Validation loss $\loss{C,C_0^\prime}$ is an unbiased estimate of loss $\loss{C,\Sigma}$ when $\loss{\cdot,\cdot}$ is \emph{additive} in its second argument so that 
 \begin{equation*}\label{eq:additivity}
 \loss{C,z_1} + \loss{C,z_2} \equiv \loss{C,z_1+z_2}
 \end{equation*}

Then the absence of bias is shown:
\begin{equation*}
    \E{\loss{C,C_0^\prime}}=\loss{C,\E{C_0^\prime}}=\loss{C,\Sigma}
\end{equation*}
} 
of loss $\loss{C,\Sigma}$.

Therefore, estimators resulting in consistently lower validation loss can be inferred to produce estimates that are closer to truth than estimators with higher validation loss.

\noindent\rule{8cm}{4pt}
